---
title: "HW3"
author: "Md Reza"
date: "10/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***

### Problem 6:
__(a)__

The equation is: 

$p(X) = \frac{e^{\beta_0+\beta_1 X_1+\beta_2 X_2}}{1+e^{\beta_0+\beta_1 X_1+\beta_2 X_2}}$

and plugging in the values we get: 

$p(X) = \frac{e^{-6 + 0.05 \times 40 + 1 \times 3.5}}{1+e^{-6+0.05 \times 40 + 1 \times 3.5}} =$


```{r}
exp(-6+0.05*40+1*3.5)/(1+exp(-6+0.05*40+1*3.5))
```
Estimated probability: `0.38`

__(b)__

The equation for predicting the probability is: 

$0.5 = \frac{e^{-6 + 0.05 X_1 + 1 \times 3.5}}{1+e^{-6+0.05 X_1 + 1 \times 3.5}}$

by taking the logarithm of both side we get:   

$log(\frac{0.5}{1-0.5}) = -6 + 0.05 X_1 + 1 \times 3.5$

```{r}
(log(0.5/(1-0.5)) + 6 - 3.5*1)/0.05
```

which indiactes that student needs to study at least `50` hours to get an _A_.


***

### Problem 7:

*By plug in the values of X and other parameters in the equation $(i.e. pk(x))$ we get:*


$p_{|yes}(4) = \frac {0.8 \exp(- \frac {1} {2 \times 36} (4 - 10)^2)} {0.8 \exp(- \frac {1} {2 \times 36} (4 - 10)^2) + (1-0.8) \exp(- \frac {1} {2 \times 36} (4 - 0)^2)}$

```{r}
(0.8*exp(-1/(2*36)*(4-10)^2))/(0.8*exp(-1/(2*36)*(4-10)^2)+(1-0.8)*exp(-1/(2*36)*(4-0)^2))
```

*The probability that a company will issue a dividend this year given that the percentage of last year reture was X=4 is: `75.2%`.*

***

### Problem 9:

__(a)__

We want to solve:

$0.37 = \frac{p{(X)}}{1-p{(X)}}$

that can be transformed into:

$\frac{1}{0.37} = \frac{1-p{(X)}}{p{(X)}} = \frac{1}{p{(X)}}-1$

Finally, we get:  

$p{(X)} = \frac{1}{\frac{1}{0.37}+1}$

```{r}
1/(1/0.37+1)
```

*So the probability of default is `27.0%`*

__(b)__

We want to solve:

$\frac{p{(X)}}{1-p{(X)}} = \frac{0.16}{1-0.16}$ = 0.19

```{r}
0.16/(1-0.16)
```

*So the odds of default is `19%`*.

***

### Problem 10:
__(a)__

```{r, warning=FALSE, message=FALSE}
require(ISLR)
data(Weekly)
summary(Weekly)
pairs(Weekly)
```

*Unlike `Smarket` data set there are positive correlations between `Year` and `Volume`.*

__(b)__

```{r, warning=FALSE, message=FALSE}
fit.logit <- glm(Direction~., data=Weekly[,c(2:7,9)], family=binomial)
summary(fit.logit)
```

*It seems `Lag2` has a statistically significant predictive value.*

__(c)__

```{r, warning=FALSE, message=FALSE}
logit.prob <- predict(fit.logit, Weekly, type="response")
logit.pred <- ifelse(logit.prob > 0.5, "Up", "Down")
table(logit.pred, Weekly$Direction)
(54+557)/nrow(Weekly)  # Accuracy=0.56

# with "Down" we get:
54/(54+48)

# with "Up" we get:
557/(430+557)
```

Therefore, Model with prediction = "Up" has the higher accuracy.

__(d)__

```{r, warning=FALSE, message=FALSE}
train.yrs <- Weekly$Year %in% (1990:2008)
train <- Weekly[train.yrs,]
test <- Weekly[!train.yrs,]
fit2 <- glm(Direction~Lag2, data=train, family=binomial)
fit2.prob <- predict(fit2, test, type="response")
fit2.pred <- ifelse(fit2.prob > 0.5, "Up", "Down")
table(fit2.pred, test$Direction)
mean(fit2.pred == test$Direction)  
```
The accuracy = 0.625

__(e)__

```{r, warning=FALSE, message=FALSE}
require(MASS)
fit.lda <- lda(Direction~Lag2, data=train)
fit.lda.pred <- predict(fit.lda, test)$class
table(fit.lda.pred, test$Direction)
mean(fit.lda.pred == test$Direction)  
```
The Accuracy = 0.625

__(f)__

```{r, warning=FALSE, message=FALSE}
fit.qda <- qda(Direction~Lag2, data=train)
fit.qda.pred <- predict(fit.qda, test)$class
table(fit.qda.pred, test$Direction)
mean(fit.qda.pred == test$Direction)  
```

The accuracy = 0.587

__(g)__

```{r, warning=FALSE, message=FALSE}
require(class)
set.seed(1)
train.X <- as.matrix(train$Lag2)
test.X <- as.matrix(test$Lag2)
knn.pred <- knn(train.X, test.X, train$Direction, k=1)
table(knn.pred, test$Direction)
mean(knn.pred == test$Direction)  
```

The accuracy = 0.500

__(h)__

*The `Logistic Regression` and `LDA` methods appears to provide the best results on this data.*


***


### Bonus Points

### Problem 12:

__(a)__

```{r}
Power <- function() {
  print(2^3)
}
Power()
```

__(b)__

```{r}
Power2 <- function(x, a) {
  print(x^a)
}
Power2(3,8)
```

__(c)__

```{r}
Power2(10,3)
Power2(8,17)
Power2(131,3)
```

__(d)__

```{r}
Power3 <- function(x , a) {
    result <- x^a
    return(result)
}
Power3(3,8)
```

__(e)__

```{r}
x <- 1:10
plot(x, Power3(x,2), log="y", main="Log of x^2 versus Log of x",
     xlab="Log of x", ylab="Log of x^2")
```

__(f)__

```{r}
PlotPower <- function(x, a) {
  plot(x, Power3(x,2), main="x^2 versus x",
       xlab="x", ylab=paste0("x^",a))
}
PlotPower(1:10,3)
```

***